---
title: "group_proj"
author: "Rachel Montgomery"
date: "2022-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

```{r}
library(readr)
UFO_and_Weather <- read_csv("UFO_and_Weather.csv", 
    col_types = cols(month = col_character(), 
        hour = col_time(format = "%H")))

View(UFO_and_Weather)

```
```{r}
#making new column with month names

UFO_and_Weather <- UFO_and_Weather %>%
  mutate(Month= case_when(
    month=="1" ~ "January",
    month=="2" ~ "February",
    month=="3" ~ "March",
    month=="4" ~ "April",
    month=="5" ~ "May",
    month=="6" ~ "June",
    month=="7" ~ "July",
    month=="8" ~ "August",
    month=="9" ~ "September",
    month=="10" ~ "October",
    month=="11" ~ "November",
    month=="12" ~ "December"))

```



Goals:
Are UFO sightings more prevalent during certain months of the year?
What time of day do UFO sightings occur? Are reports less reliable in these times (e.g. Friday/Saturday night, low vision)
How long does the average UFO sighting last?

```{r}
#Understanding what the data represents
View(UFO_and_Weather)

head(UFO_and_Weather)

dim(UFO_and_Weather)

str(UFO_and_Weather)
```
22,482 obs and 18 variables. 

#1 Are UFO sightings more prevalent during certain months of the year?
```{r}
ggplot(UFO_and_Weather, aes(x=Month))+
  geom_bar()+
  coord_flip()

#July has the most 
```


# 1.2 Is there a most popular season for UFO sightings?

```{r}
#first creating a season variable
UFO_and_Weather <- UFO_and_Weather %>%
  mutate(season= case_when(
    month=="12" | month=="1" | month=="2" ~ "Winter"  ,
    month=="3" |  month=="4" | month=="5" ~ "Spring",
    month=="6" | month=="7" | month=="8" ~ "Summer",
    month=="9" | month=="10" |  month=="11" ~ "Fall"))

#spring is march - may (3,4,5)
#summer is june to august (6,7,8)
#fall is september to november (9,10,11)
#winter is december to february (12, 1, 2)
```



#2 What time of day do UFO sightings occur?
```{r}
#first making time into am and pm hour times
UFO_and_Weather <- UFO_and_Weather %>%
  mutate(time_of_day= case_when(
    hour >=6 & hour <= 18 ~ "day",
    hour < 5  ~ "night",
    hour >18 ~ "night"))

#next, making day and night variables
#day is 6 am to 6 pm (6  to 18)
#night is 12 am to 6am, and 6 pm to 12 pm (0-6, and 18-24)
UFO_and_Weather <- UFO_and_Weather %>%
  mutate(time_of_day= case_when(
    hour >=6 & hour <= 18 ~ "day",
    hour < 5  ~ "night",
    hour >18 ~ "night"))

```


```{r}
ggplot(UFO_and_Weather, aes(x=hour))+
  geom_bar()+
  coord_flip()
```
We can see that the most sightings are at night, specifically between the hours of 6 pm and 11 pm.


```{r}
ggplot(UFO_and_Weather, aes(x =month)) +
        geom_boxplot() +
        scale_x_discrete(name = "Month") +
        scale_y_continuous(name = "Mean ozone in\nparts per billion") + 
        ggtitle("Boxplot of mean ozone by month") +
        theme_minimal()



ggplot(air_sub, aes(x = month_cat, y = Ozone)) +
        geom_boxplot() +
        scale_x_discrete(name = "Month") +
        scale_y_continuous(name = "Mean ozone in\nparts per billion",
                          breaks = seq(0, 175, 25),
                          limits=c(0, 175)) + 
        ggtitle("Boxplot of mean ozone by month") +
        theme_minimal()
```



```{r}
ggplot(UFO_and_Weather, aes(x=season))+
  geom_bar()+
  coord_flip()
```
We can see that the fall months have the most sightings, with summer having the 2nd most.


#3 How long does the average UFO sighting last?
We can't do this because the duration variable was taken out.










