---
title: "Findings for text analysis and frequency of sightings"
author: "Rachel Montgomery"
date: "2022-11-29"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(tinytex.verbose = TRUE)
```

# Findings in UFO Project

This rmd file encompasses the exploratory data analysis regarding the following areas of interest:

-   Frequency of sightings

    -   Exploring if sightings are more prevalent during certain months of the year

    -   If seasons impact the frequency of sightings

    -   Exploring which times of day are the most popular for sightings

        -   Investigating if reports are less reliable in these times (e.g. Friday/Saturday night, low vision)
        -   If the time of day impacts what shapes of UFOs are seen

    -   Which days of the week are the most common for sightings

    -   If the frequency of UFO sightings changed over the years?

    -   Which states have the most sightings

-   Text analysis

    -   In reports, which words are the most common

    -   Creating word clouds generated from reports

    -   Sentiment Analysis

## Set up: loading in data and libraries

Loading in needed libraries:

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidytext)
library(textdata)
library(tm)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
library(syuzhet)

library(gridExtra)
library(patchwork)
library(factoextra)
```

Reading in data:

```{r}
library(readr)
UFO_and_Weather <- read_csv("UFO_and_Weather.csv", 
    col_types = cols(month = col_character(), 
        hour = col_time(format = "%H")))

#taking out non-us states
UFO_and_Weather <- subset(UFO_and_Weather, !(state %in% c("BC", "QC", "ON", "NS", "AB", "MB", "SK", "YT", "NL")))
```

### Viewing Data

```{r}
#Understanding what the data represents
View(UFO_and_Weather)

head(UFO_and_Weather)

dim(UFO_and_Weather)

str(UFO_and_Weather)
```

## Data Analysis


### Frequency of Sightings

#### Which season has the most sightings? Has this changed over the years?

```{r}
#first creating a season variable
UFO_and_Weather <- UFO_and_Weather %>%
  mutate(season= case_when(
    month=="12" | month=="1" | month=="2" ~ "Winter"  ,
    month=="3" |  month=="4" | month=="5" ~ "Spring",
    month=="6" | month=="7" | month=="8" ~ "Summer",
    month=="9" | month=="10" |  month=="11" ~ "Fall"))

#spring is march - may (3,4,5)
#summer is june to august (6,7,8)
#fall is september to november (9,10,11)
#winter is december to february (12, 1, 2)
```

```{r}
ggplot(UFO_and_Weather, aes(x=season))+
  geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "UFO Sightings grouped by the seasons") +
  xlab("Season") +
  ylab("Number of sightings") +
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),) +
  coord_flip()

```

We can see that the fall months have the most sightings, with summer having the 2nd most.

#### Which months have the highest count of sightings? Has this changed over the years?

```{r}
#first making new column with month names

UFO_and_Weather <- UFO_and_Weather %>%
  mutate(Month= case_when(
    month=="1" ~ "January",
    month=="2" ~ "February",
    month=="3" ~ "March",
    month=="4" ~ "April",
    month=="5" ~ "May",
    month=="6" ~ "June",
    month=="7" ~ "July",
    month=="8" ~ "August",
    month=="9" ~ "September",
    month=="10" ~ "October",
    month=="11" ~ "November",
    month=="12" ~ "December"))

#plotting 
ggplot(UFO_and_Weather, aes(x = reorder(Month, Month, FUN = length))) +
  geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "UFO Sightings thoughout the year") +
  xlab("Month") +
  ylab("Number of sightings") +
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),
        ) +
  coord_flip()
```

We can see that July has the most by quite a good margin. When we examine year by year, for 4/5 years July has either the most or the 2nd most sightings. This has me think that we should plot the number of sightings per each day in July. Maybe this is because of the 4th of july - a lot of fireworks and air shows happening? Because its summer more people are out and about and looking at the sky?

Let's investigate if how the popular months vary by the years.

```{r}
#let's see if this pattern of mostly being in the fall is a yearly pattern!

#UFO_and_Weather$Month = factor(UFO_and_Weather$Month, levels=month.abb)

weather_2019 <- UFO_and_Weather %>%
    filter(year == 2019)
#plotting
w19 <- ggplot(weather_2019, aes(x=reorder(Month, Month, FUN=length)))+
   geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "2019") +
  xlab("Month") +
  ylab("Count") +
   theme_minimal() +
  theme(plot.title = element_text(size = 10, face = "bold",hjust = 0.5),
        text = element_text(size = 5),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 6),
        ) +
  scale_x_discrete(limits = month.abb) +
  coord_flip()

weather_2018 <- UFO_and_Weather %>%
    filter(year == 2018)
#plotting
w18 <- ggplot(weather_2018, aes(x=reorder(Month, Month, FUN=length)))+
   geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "2018") +
  xlab("Month") +
  ylab("Count") +
    theme_minimal() +
  theme(plot.title = element_text(size = 10, face = "bold",hjust = 0.5),
        text = element_text(size = 5),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 6),
        ) +
  coord_flip()

weather_2017 <- UFO_and_Weather %>%
    filter(year == 2017)
#plotting
w17 <- ggplot(weather_2017, aes(x=reorder(Month, Month, FUN=length)))+
   geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "2017") +
  xlab("Month") +
  ylab("Count") +
  theme_minimal() +
  theme(plot.title = element_text(size = 10, face = "bold",hjust = 0.5),
        text = element_text(size = 5),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 6),
        ) +
  coord_flip()

weather_2016 <- UFO_and_Weather %>%
    filter(year == 2016)
#plotting
w16 <- ggplot(weather_2016, aes(x=reorder(Month, Month, FUN=length)))+
   geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "2016") +
  xlab("Month") +
  ylab("Count") +
   theme_minimal() +
  theme(plot.title = element_text(size = 10, face = "bold",hjust = 0.5),
        text = element_text(size = 5),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 6),
        ) +
  coord_flip()

weather_2015 <- UFO_and_Weather %>%
    filter(year == 2015)
#plotting
w15 <- ggplot(weather_2015, aes(x=reorder(Month, Month, FUN=length)))+
   geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "2015") +
  xlab("Month") +
  ylab("Count") +
  theme_minimal() +
  theme(plot.title = element_text(size = 10, face = "bold",hjust = 0.5),
        text = element_text(size = 5),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 6),
        ) +
  coord_flip()


grid.arrange(w15,w16,w17,w18,w19, nrow = 2, top="Sightings per Month")#

```


We can consistently see that July has a large amount of sightings. Let's investigate just the month of July.

For the month of July, which days are the most popular?

```{r}
just_july <- UFO_and_Weather %>%
  filter(month== 7) 

#plotting
ggplot(just_july, aes(x = reorder(day, day, FUN = length))) +
  geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "UFO Sightings in the month of July") +
  xlab("Day") +
  ylab("Number of sightings") +
   theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 7)) +
        coord_flip()
```

Wow!!! The fourth of july has about 2.5x more than all other days!! Some possible explanations for this are because there are a ton of fireworks going off, which cause light and sound to be in the sky, and could be mistaken for UFOS. Also there are air shows that occur, which could also be reported as UFOS. Because of these festivities, many people are out at night and looking at the sky, so mere exposure effect could be play. Additionally, it should be noted that this is a national holiday, and one that is often celebrated with alcohol, which conflicts the reliability of these reports.

Overall, the 4th of July is an outlier and should be treated as such in following analysis.


#### Which times of the day are the most popular for sightings?

```{r}
#first making time into am and pm hour times
UFO_and_Weather <- UFO_and_Weather %>%
  mutate(time_of_day= case_when(
    hour >=6 & hour <= 18 ~ "day",
    hour < 5  ~ "night",
    hour >18 ~ "night"))

#next, making day and night variables
#day is 6 am to 6 pm (6  to 18)
#night is 12 am to 6am, and 6 pm to 12 pm (0-6, and 18-24)
UFO_and_Weather <- UFO_and_Weather %>%
  mutate(time_of_day= case_when(
    hour >=6 & hour <= 18 ~ "day",
    hour < 5  ~ "night",
    hour >18 ~ "night"))

#now plotting
ggplot(UFO_and_Weather, aes(x=hour)) + 
  geom_bar(alpha = 0.9,position=position_dodge(width=0.8) ,fill = "darkslateblue") +
  scale_fill_brewer(palette = "Greens") +
                labs(title="UFO Sightings throughout the day") +
                xlab("Hour of the day") +
                ylab("Number of sightings") +
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8))
 
```

We can see that the most sightings are at night, specifically between the hours of 6 pm and 11 pm. The histogram clearly shows that the majority of the sightings occur when there is no or very little light. However, it's worth noting that there have been reports of UFOs during the day as well.

##### Is there a relationship between the UFO shapes and time?

We can look at the relationship between the shapes and the time of day. This may explain why the light shape is the most frequently reported shape.

```{r}
#subsetting data for just shapes and time
shapesDaytime <- 
  UFO_and_Weather %>% 
  group_by(hour, shape) %>% 
  summarize(count=n());

#plotting 
ggplot(shapesDaytime, aes(x=hour, y=shape)) + 
  geom_point(aes(color=count, size=count, )) + 
  scale_colour_gradient(low = "lightblue", high="darkslateblue") +
  labs(x = "Hour of the day", y = "UFO Shape", 
       title="Correlation between daytime / UFO Shape", 
       subtitle = "Sightings during the day") +
theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 7))
```

We can see from the plot that the shapes are more prevalent/persistent at night as well. We can see that light appears more frequently at night and in the evening, but less frequently during the day than other common shapes.

#### What days of the week are the most popular for sightings?

```{r}
#To see the most popular day of the week, we need to manipulate the original date time variable
UFO_and_Weather$weekday <- weekdays(UFO_and_Weather$date_time)

#plotting
ggplot(UFO_and_Weather, aes(x=reorder(weekday, weekday, FUN=length)))+
 geom_bar(alpha = 0.9,  width=0.7, position=position_dodge(width=0.8), fill ="darkslateblue" ) +
  labs(title = "Sightings by day of week") +
  xlab("Day of Week") +
  ylab("Count") +
   theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 7)) +
        coord_flip()
```

Weekends have the most sightings, which is interesting. Combining this with the fact that most are night, infers that many sightings are during "drinking" or social hours at night. This could hinder the reliability of sightings.

#### Has the frequency of UFO sightings changed over the years?

```{r}
#Sightings by year
sightingsYear <- 
  UFO_and_Weather %>% group_by(year) %>% 
  summarize(count=n());

#getting the reports by year
reportsYear <- 
  UFO_and_Weather %>% group_by(year) %>% 
  summarize(count=n())

#plotting
ggplot(sightingsYear, aes(x=year, y=count))  + 
  geom_smooth(method="lm") +
  labs(x = "Year", y = "red= sightings", 
       title="UFO sightings / UFO reports by year", 
       subtitle = "Sightings during the day") +
    theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 7)) 

```


#### Which states have the most sightings?

```{r}
ggplot(UFO_and_Weather, aes(x=reorder(state, state, FUN=length))) + 
  stat_count(fill="darkslateblue") +
  labs(x = "State", y = "Number of sightings", 
       title="UFO sightings in United States", 
       subtitle = "Sightings by state")+
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(angle=45, size=4, hjust=1,face = "bold"),
        axis.text.y = element_text(size = 7))

```

The states with the most reports are California, Florida, and Washington.

However, an important point to note is that, based on population density, the state of Washington has a higher density of sightings than California. California and Washington have both legalized marijuana for recreational use, which could explain why there have been so many sightings on the west coast.

#### Which cities have the most sightings?
```{r}
#city state counts of sightings
state_counts <- UFO_and_Weather %>%
  filter(state != '')%>%
  count(city, state, sort=TRUE)%>%
  unite('location', -n, sep=',')


#visualize with lollipop chart
state_counts %>%
  filter(n > 90)%>%
  mutate(x = factor(location))%>%
  ggplot(aes(x, n))+
  geom_segment(aes(x=reorder(x,n), xend=x, y=0, yend=n), size=1)+
  geom_point(color='darkslateblue', size=2, alpha=.6)+
  theme(axis.text.y = element_text(size = 7))+
  labs(title="Which cities have had the most sightings?") +
  xlab("Count") +
  ylab("Cities")+
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 7)) +
        coord_flip()
```


### Text Analysis

In this section, we will perform text analysis on the text from reports.

#### Word Frequency

```{r}
#Step 1: Tokenize corpus

words <- UFO_and_Weather %>%
  select(text) %>%
  unnest_tokens(word, text)

head(words)

#Now, we'll generate a count of the words, sort by the number of times the word occurs, and then plot the top 15 words in a bar plot

words %>% count(word, sort = T) %>% slice(1:15) %>%
  ggplot(aes(x = reorder(word, n, function(n) -n), y = n)) +
  geom_bar(stat = 'identity' ,fill = "darkslateblue") +
  #theme_light() +
  #theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("Words") +
  ylab("Count") +
  ggtitle(" Word Count (with stop words)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8))
```

As we can see, the most popular words (at the moment) are words that are commonly used in the English language, regardless of context. These are called stop words. Examples of stop words in English are "a", "the", "is", "are" and etc.

In the next step, we'll remove the stop words so we can focus on the important words instead.

```{r}
#Step 2: Using the `TidyText` package, remove stop words and generate a new word count
ufo_no_stop <- words %>% 
  anti_join(stop_words)

#creating personal stop words to take out
#these words either have to do with reporting, or wern't topic words
personal_stop_words <- stop_words %>% 
  dplyr::select(-lexicon) %>%
  bind_rows(data.frame(word = c("nuforc", "note", "time", "looked", "appeared","minutes")))

ufo_no_stop <- ufo_no_stop %>% 
  anti_join(personal_stop_words)

#plotting
ufo_no_stop %>%
  count(word, sort = T) %>% 
  slice(1:15) %>% 
  ggplot(aes(x = reorder(word, n, function(n) -n), y = n)) + 
  geom_bar(stat = "identity", alpha = 0.9, fill = "darkslateblue") + 
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(angle=45,size = 9)) +
  labs(title ="Word Frequency without Stop Words" )+
  xlab("Words") +
  ylab("Count")

```

We can see the most common words are those typical for a UFO report. Light, sky, object, moving, and looked all make sense here.

#### Word Clouds

A word cloud is a collection, or cluster, of words depicted in different sizes. The bigger and bolder the word appears, the more often it's mentioned within a given text and the more important it is.

First, we'll build a document matrix, which is a table containing the frequency of the words. Column names are words and row names are documents.

```{r}
#Creating a term-document matrix

tdm <- TermDocumentMatrix(ufo_no_stop)
m <- as.matrix(tdm)
v <- sort(rowSums(m),decreasing=TRUE)
term_dox_matrix <- data.frame(word = names(v),freq=v)
head(term_dox_matrix, 10)
```

We can see that the most common words are light, lights, sky, object, and bright. This follows what we plotted earlier.

Now, we can create a word cloud.

```{r}
#generate word cloud
set.seed(1234)
wordcloud(words = term_dox_matrix $word, freq = term_dox_matrix$freq, min.freq = 1, 
          colors=brewer.pal(8, "Dark2"),  scale=c(3.5,0.25))

#In order to have words not cropped out from spacing issues, changed scale

wordlcoud(words = term_dox_matrix $word, freq = term_dox_matrix$freq)


set.seed(1234) # for reproducibility 
wordcloud(words =term_dox_matrix $word, freq = term_dox_matrix$freq, min.freq = 1,           max.words=200, random.order=FALSE, rot.per=0.35,            colors=brewer.pal(8, "Dark2"))
```

```{r}
#We can also create word clouds in specific shapes

wordcloud2(data=term_dox_matrix , size = 0.4, shape = 'diamond')


wordcloud2(term_dox_matrix , color = "random-light", backgroundColor = "grey")

```

#### Sentiment Analysis

Let's explore sentiment analysis.

We'll be using the Syuzhet package for generating sentiment scores, which has four sentiment dictionaries and offers a method for accessing the sentiment extraction tool developed in the NLP group at Stanford.

```{r}
emotions <- get_nrc_sentiment(ufo_no_stop$word)
head(emotions)
```

```{r}
barplot(colSums(emotions),
        las = 2,
        col = rainbow(10),
        ylab = 'Count',
        main = 'Sentiment Scores')
```

```{r}
sentiments <- get_sentiments("nrc")

df_sentiments1 <- ufo_no_stop %>% left_join(sentiments)

df_sentiments_filtered1 <- df_sentiments1 %>% 
  filter(!is.na(sentiment)) %>% 
  group_by(sentiment) %>% 
  summarize(n = n())

df_sentiments_filtered1 %>% 
  ggplot(aes(x = reorder(sentiment, n, function(n) -n), y = n)) + 
  geom_bar(stat = "identity", alpha=0.7, fill="forestgreen") + 
  theme_light() +
  theme(plot.title = element_text(size = 12,
      family = "sans",
      face = "bold",hjust = 0.5),
      panel.background = element_rect(fill = "grey88",
                                colour = "grey88",
                                size = 0.5, linetype = "solid"),
      panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "grey92"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "grey92"),
    text = element_text(size = 11, family = "sans"),
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 60, hjust = 1,  size=10),
    axis.text.y = element_text(size = 9)) + 
    labs(title ="Sentiments in UFO Reports")+
  xlab("Emotions") +
  ylab("Count")
```


The most common sentiment by a large margin is positive.

Overall, this sentiment analysis isn't as relevant and is subjective, and therefore will not be included in the final report of findings.
