---
title: "Missingness and Data Validations"
author: "Yuning Wu"
date: "2022-12-11"
output: html_document
---

# Libraries

```{r}
library(tidyverse)
library(VIM)
library(naniar)
library(gridExtra)
```

# Load Data

```{r}
ufo_uncleaned <- read_csv("../01_Preprocessing/nuforc_reports.csv")
```

# Data Cleaning and Validation

## Visualize missingness:

```{r}
x <- gg_miss_var(ufo_uncleaned)+
  ggtitle("Total Missingness in All Variables")
x
ggsave("missing1.png")
```

```{r}
y <- gg_miss_var(ufo_uncleaned, show_pct = TRUE)+
  ggtitle("Total Percentage Missingness in All Variables")
y
ggsave("missing2.png")
```

### Matrix Plot
```{r}
matrixplot(ufo_uncleaned)
ggsave("missing3.png")
```

- No apparent patterns shown in the matrix plot.

## Investigate the top most missingness:

### Date_time and posted

```{r}
# sort by date_time
matrixplot(ufo_uncleaned, sortby = 4)

sum(is.na(ufo_uncleaned$date_time))
sum(is.na(ufo_uncleaned$posted))

ufo_uncleaned %>% 
  filter(is.na(date_time)) %>% 
  select(posted) %>% 
  summarise(num_na_posted = sum(is.na(posted)))
```

- Missingness in date_time and posted is highly correlated. In fact, there are 1187 missing values in each and they are from the same observations.

### state and city

```{r}
# sort by state
matrixplot(ufo_uncleaned, sortby = 3)

# sort by city
matrixplot(ufo_uncleaned, sortby = 2)

ufo_uncleaned %>% 
  filter(is.na(state)) %>% 
  select(city)
```

- Missingness in state is only correlated with missingness in city_latitude and city_longitude which is expected.  
- Cities that are missing state data are mostly foreign cities (the scope of our project is within the US, so we should remove these).


```{r}
sum(is.na(ufo_uncleaned$city_latitude))
sum(is.na(ufo_uncleaned$city_longitude))

ufo_uncleaned_1 <- ufo_uncleaned %>% 
  arrange(date_time)

matrixplot(ufo_uncleaned_1)
```

- There are many NAs in longitude and latitude, however they are always missing together.  
- The missingness in these 2 variables doesn't seem to correlated to any other variable.  
- Clear to drop, but since there are too many, we decided to keep two versions one with no NA values (ufo), one with NA values in city_longitude and city_latitude (ufo_extended).

### duration

```{r}
matrixplot(ufo_uncleaned, sortby = 6)
```

- No correlation, clear to drop.

### shape

```{r}
matrixplot(ufo_uncleaned, sortby = 5)
```

- No correlation, clear to drop.

### Dropping all NAs from the df to create "ufo"

```{r}
ufo <- na.omit(ufo_uncleaned)
sum(is.na(ufo))
```

### Dropping all NAs except the ones from city_longitude and city_latitude to create "ufo_extended"

```{r}
ufo_extended <- ufo_uncleaned %>% 
  filter(!is.na(state)) %>% 
  filter(!is.na(city)) %>% 
  filter(!is.na(duration)) %>% 
  filter(!is.na(shape)) %>% 
  filter(!is.na(date_time)) %>% 
  filter(!is.na(posted)) %>% 
  filter(!is.na(text)) %>% 
  filter(!is.na(stats)) %>% 
  filter(!is.na(summary))
```

```{r}
sum(is.na(ufo_extended))
gg_miss_var(ufo_extended)
```

# Data Validations
```{r}
ufo %>% 
  chain_start %>% 
  # check if ufo contains only valid US states
  assert(in_set(state.abb), state) %>%
  # check if latitude is valid
  assert(within_bounds(-90, 90), city_latitude) %>% 
  # check if longitude is valid
  assert(within_bounds(-180, 180), city_longitude) %>% 
  chain_end(error_fun = warn_report)
```

- Discovery: There are some non-US states in the dataset. We need to remove them before our analyses on UFO sightings in the US.
