---
title: "Population, Shapes, Phenomena"
author: "Edmund Hui, Rio Jia, Rachel Montgomery, Yuning Wu"
date: "2022-11-02"
output: pdf_document
---

# Dependencies & Read in Data

```{r}
#install.packages("tidyverse")
#install.packages("stringr")
#install.packages("usmap")
#install.packages("cowplot")
#install.packages("wordcloud")
library(wordcloud)
library(cowplot)
library(tidyverse)
library(stringr)
library(usmap)
require(ggplot2)
require(ggmap)
require(maps)
```

```{r}
df <- read_csv("../02_Data/Clean_UFO_Dataset.Rmd")
df
```

# Question 1: Do UFO sightings happen in more densely populated areas?

We would have to add in some sort of population/census data, but could be interesting to look into

```{r}
citypop <- read_csv("../01_Preprocessing/populations_by_city.csv")

# Filter only the municipalities
citypop <- citypop %>% filter(TYPE %in% c("city", "town", "village")) 
citypop
```

```{r}
df <- df %>% 
  left_join(citypop, by = c("city"="SHORTNAME", "state"="STSHORT")) %>% 
  rename(population_2021 = POPESTIMATE2021, geo_class=TYPE)

df 
```

## Group by municipalities (administrative division)

```{r}
muni <- df %>% drop_na(geo_class) %>% group_by(geo_class) %>% summarise(count=n()) 
muni %>% ggplot(aes(x=geo_class, y=count)) + geom_col() + 
  theme_minimal() + 
  labs(title="Number of UFO Sightings by Administrative Type") + 
  xlab("Administrative Division") + 
  ylab("Number of Sightings")
```
## Group By Population

Use case_when to split population into even & logical levels

```{r}
#mean(citypop$POPESTIMATE2021)
#mean(df$population_2021, na.rm=TRUE)
```

## UFO Reports Per Capita / Per State 

```{r}
states <- read_csv("../01_Preprocessing/statepop.csv", col_names=FALSE)
states <- states %>% select(X1, X4)
names(states) <- c("state", "population_2021")
states$state <- substr(states$state, 2, 100)
states$state <- state.abb[match(states$state,state.name)]
states <- states %>% drop_na()
```
```{r}
sight_counts <- df %>% group_by(state) %>% summarise(count=n())
states <- states %>% left_join(sight_counts, by="state") %>% mutate(obs_100k = (count/population_2021)*100000)
```

```{r}
states %>% arrange(obs_100k)
```

```{r}
plot_usmap(data = states, values = "count", color = "black") + 
  scale_fill_continuous( low = "white", high = "red", name = "Population (2015)", label = scales::comma) + 
  theme(legend.position = "right") + 
  labs(title="UFO Sightings Per 100K Population by State", subtitle="Montana has the highest sightings per capita with 20.19 sightings per 100k Population")
```


```{r}
plot_usmap(data = states, values = "obs_100k", color = "black") + 
  scale_fill_continuous( low = "white", high = "red", name = "Population (2015)", label = scales::comma) + 
  theme(legend.position = "right") + 
  labs(title="UFO Sightings Per 100K Population by State", subtitle="Montana has the highest sightings per capita with 20.19 sightings per 100k Population")
```


# Question 2: What are the most common UFO descriptions?

## General Shape Analysis

```{r}
df %>% drop_na(shape) %>% filter(shape %in% c("light", "circle", "triangle", "unknown", "fireball")) %>%
  ggplot() + 
  geom_bar(aes(x = reorder(shape, shape, function(x)-length(x))), fill="#D8DADC") +  
  #scale_x_discrete(guide = guide_axis(angle = 90)) + 
  labs(title="Most Common UFO Shapes Sighted", subtitle="Only the top five shapes displayed") +
  xlab("Shape of UFO Sighted") + 
  ylab("Number of Sightings") + 
  theme_minimal() 

ggsave("shapes_overall.png")
```


```{r}
df %>% drop_na(state) %>% 
  filter(state %in% c("CA", "FL", "WA", "AZ", "NY")) %>%
  ggplot() + 
  geom_bar(aes(x = reorder(state, state, function(x)-length(x))), fill="#D8DADC") +  
  labs(title="States with the Most UFO Sightings", subtitle="Only the top five states displayed") +
  xlab("State") + 
  ylab("Number of Sightings") + 
  theme_minimal() 

ggsave("states_overall.png")
```

It is quite likely that "light" shapes occur at night, where it could have very likely been a mistaken street light, plane, or something else. This is corroborated by the time distribution of "light" observations. 

```{r}
df %>% filter(shape=="light") %>% ggplot() + geom_bar(aes(x=hour))
```


## Which Shapes are Most Common in Each State?

```{r}
df %>% 
  group_by(state, shape) %>% 
  summarise(count=n()) %>% 
  group_by(state) %>% 
  top_n(1, count)
```


# Question 3: Do certain cultural phenomena influence UFO sightings?

--- We can add in cultural data like # of sci fi movies released in a year and see if there is a correlation, if a war is happening, etc

## 3.1 Investigate the relationships between google trends data and UFO sightings

```{r}
df_counts
```


```{r}
# Read in google trend data
trends <- read_csv("../01_Preprocessing/multiTimeline.csv", skip=1)

# Rename columns 
trends <- trends %>% rename(month=Month,ufo=`ufo: (United States)`, alien=`alien: (United States)`)

# Take only the years we need
trends <- trends %>% filter(substr(month, 1,4) %in% c("2015","2016","2017","2018","2019"))

# Count occurances each month 
df_counts <- df %>% group_by(year, month) %>% summarise(count=n()) 

# Add counts to the trends dataframe
trends$count <- df_counts$count
```

```{r}
# Min-Max scale as we are only interested in relative movements
trends$ufo_scaled <- (trends$ufo-min(trends$ufo))/(max(trends$ufo)-min(trends$ufo))
trends$alien_scaled <- (trends$alien-min(trends$alien))/(max(trends$alien)-min(trends$alien))
trends$count_scaled <- (trends$count-min(trends$count))/(max(trends$count)-min(trends$count))
```

```{r}
skip_four <- function(x) x[seq_along(x) %% 6 == 0]

trends %>% ggplot(aes(x=month, y=ufo_scaled, group=1)) + 
  geom_line(aes(colour="\"UFO\" Search Query"), alpha=0.8, linewidth=0.7) + 
  geom_line(aes(y=alien_scaled, colour="\"Alien\" Search Query"), alpha=0.7, linewidth=0.7) + 
  geom_line(aes(y=count_scaled, colour="UFO Sightings"), linewidth=1) + 
  labs(x = "Date (Year-Month)",
       y = "Scaled Observation Frequncy",
       color = "Legend") + 
  scale_colour_manual("Scaled Frequencies", 
                      breaks = c("\"UFO\" Search Query", "\"Alien\" Search Query", "UFO Sightings"),
                      values = c("red", "blue", "black")) + 
  labs(title="UFO Sightings Relative to Related Search Queries") +
  scale_x_discrete(breaks=skip_four, guide = guide_axis(angle = 45)) + 
  #scale_x_discrete(breaks=pretty_breaks(n=3)) + 
  annotate("text", x=15, y=0.98, label= "Nov \n 2015") + 
  geom_point(x=11, y=1, colour="red", shape=17, size=2) +
  theme_minimal()+
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8)) 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

ggsave("searches.jpeg")
```


## 3.1 Investigate the relationships between sci-fi movie data and ufo sightings

Data adapted from https://www.the-numbers.com/market/creative-type/Science-Fiction 

```{r}
trends %>% ggplot(aes(x=month, y=count, group=1)) + geom_line() +
  scale_x_discrete(guide = guide_axis(angle = 90)) + geom_segment(aes(x = 12, y = 400, xend = 12, yend = 700),
                  arrow = arrow(length = unit(0.3, "cm")), colour="darkslateblue") +
  
  scale_x_discrete(guide = guide_axis(angle = 90)) + geom_segment(aes(x = 24.5, y = 1300, xend = 24.5, yend = 600),
                  arrow = arrow(length = unit(0.3, "cm")), colour="darkslateblue") +
  
  scale_x_discrete(guide = guide_axis(angle = 90)) + geom_segment(aes(x = 36, y = 1000, xend = 36, yend = 900),
                  arrow = arrow(length = unit(0.3, "cm")), colour="darkslateblue") +
  
  scale_x_discrete(guide = guide_axis(angle = 90)) + geom_segment(aes(x = 41.5, y = 800, xend = 41.5, yend = 430),
                  arrow = arrow(length = unit(0.3, "cm")), colour="darkslateblue") +
  
  annotate("text", x=12, y=320, label="Rogue One: A Star Wars Story Released \n Grossing $741M", size=3) + 
  annotate("text", x=27, y=1400, label="Star Wars Ep. VII Released \n Grossing $425M", size=3) + 
  annotate("text", x=36, y=1100, label="Star Wars Ep. VIII Released \n Grossing $425M", size=3) + 
  annotate("text", x=44, y=900, label="Jurassic World \n Fallen Kingdom Released \n Grossing $390M", size=3) +
  
  xlab("Date (Year-Month)") + 
  ylab("Number of Sightings")+
  
  scale_x_discrete(breaks=skip_four, guide = guide_axis(angle = 45)) +
  labs(title="Sighting Frequencies and Major Movie Releases") + 
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8)) 
  
  
ggsave("movie_releases.jpeg")
```

```{r}
year <- c("2015", "2016", "2017", "2018", "2019")
n_movies <- c(41, 37, 40, 40, 24)
total_gross <- c(3149106114, 1527475428, 1867466016, 1939884926, 1107000236)
top_movie <- c("Star Wars Ep. VII: The Force Awakens", "Rogue One: A Star Wars Story", "Star Wars Ep. VIII: The Last Jedi", "Jurassic World: Fallen Kingdom", "Star Wars: The Rise of Skywalker")
top_movie_gross <- c(742208942, 424987707, 517218368, 417719760, 390706234)

yearly_gross <- data.frame(year, n_movies, gross, top_movie, top_movie_gross)
yearly_gross
```


```{r}
trends %>% 
  separate(month, c("year", "month"), "-") %>%
  group_by(year) %>% 
  summarise(sightings = sum(count))

yearly_gross
```


```{r}
trend_plot <- trends %>% 
  separate(month, c("year", "month"), "-") %>%
  group_by(year) %>% 
  summarise(sightings = sum(count)) %>% 
  ggplot() +
    geom_line(aes(x=year, y=sightings), group=1) + labs(title="UFO Sightings by Year") + 
    xlab("") + 
    ylab("Sightings") + 
    theme_bw()

gross_plot <- yearly_gross %>% 
  ggplot(aes(x=year, y=gross)) + 
    geom_line(group=1, , colour="red") + 
    labs(title="Gross Revenue of Sci-Fi Movies by Year") + 
  xlab("") + 
    ylab("Gross Revenue ($)") + 
    theme_bw()

n_movies_plot <- yearly_gross %>% 
  ggplot(aes(x=year, y=n_movies), ) + geom_line(group=1, colour="blue") + 
  xlab("Year") + 
    ylab("Number of Movies Released") +
    labs(title="Number of Scifi Movies Released by Year") + 
    theme_bw()

```

```{r, fig.height=10, fig.width=8}

title <- ggdraw() + 
  draw_label(
    "Miles per gallon decline with displacement and horsepower",
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(trend_plot, gross_plot, n_movies_plot, 
          align='v', 
          labels=c("A", "B", "C"),
          nrow=3 
          )

ggsave("movies_gross.jpeg")
```

```{r}
df %>% filter(!grepl("MADAR", text))
```

## 4. Investigate the relationship between Alcohol consumption per state and sightings

```{r}
alcohol <- read_csv("../01_Preprocessing/alcohol.csv")
alcohol$state <- state.abb[match(alcohol$state,state.name)]
alcohol
```

```{r}
alc <- states %>% 
  left_join(alcohol, by="state")
```

```{r}
cor(alc$obs_100k, alc$gallonspercap)
```


```{r}
sd(alc$gallonspercap)
```


```{r}
states %>% 
  left_join(alcohol, by="state") %>% 
  ggplot() + 
    geom_point(aes(x=gallonspercap, y=obs_100k)) + 
    geom_text(aes(x=gallonspercap, y=obs_100k, label=ifelse(rank %in% c(1, 51), state, ""), vjust=2)) + 
    geom_vline(xintercept=2.4634, linetype="dashed", color="darkslateblue") + 
    geom_text(aes(x=2.4634, label="Average Consumption", y=18), colour="darkslateblue", angle=90, size=3, vjust=-1, hjust=0.6) + 
    labs(title="Observations per 100K against Alcohol Consumption For Each State") + 
    xlab("Alcohol Consumption (Gallons Per Capita)") + 
    ylab("UFO Observations (Per 100K Population)") + 
    theme_minimal() + 
    theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8))

ggsave("alcohol_consumption.jpeg")
```

## 5. Educational Attainment and UFO SIghtings

```{r}
education <- read_csv("../01_Preprocessing/education.csv")
education$state <- state.abb[match(education$state,state.name)]
```

```{r}
edu <- states %>% 
  left_join(education, by="state")
```


```{r}
cor(edu$obs_100k, edu$PercentHighSchoolOrHigher)
max_ed <- max(edu$PercentHighSchoolOrHigher)
min_ed <- min(edu$PercentHighSchoolOrHigher)
mean_ed <- mean(edu$PercentHighSchoolOrHigher)
```

```{r}
states %>% 
  left_join(education, by="state") %>% ggplot() + 
  geom_point(aes(x=PercentHighSchoolOrHigher, y=obs_100k)) + 
  geom_text(aes(x=PercentHighSchoolOrHigher, y=obs_100k, label=ifelse(PercentHighSchoolOrHigher > 93.3 | PercentHighSchoolOrHigher < 85 , state, ""), vjust=-1)) +
  geom_vline(xintercept=mean_ed, linetype="dashed", color="darkslateblue") + 
  geom_text(aes(x=mean_ed, label="Average Graduation Rate", y=18), colour="darkslateblue", angle=90, size=3, vjust=-1, hjust=0.6) + 
  labs(title="Observations per 100K against Percentage Graduating High School") + 
  xlab("Percentage of Population Graduating High School") + 
  ylab("UFO Observations (Per 100K Population)") + 
  scale_y_continuous(limits = c(0, 22)) + 
 
  theme_minimal() +
  
    theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8))

ggsave("graduating.jpeg")
```

```{r}
intel <- read_csv("../01_Preprocessing/iq.csv")
intel$state <- state.abb[match(intel$state,state.name)]
```

```{r}
states %>% 
  left_join(intel, by="state") %>% ggplot() + 
  geom_point(aes(x=averageIQ, y=obs_100k)) + 
  labs(title="Observations per 100K against Intelligence Quotient For Each State") + 
  xlab("Average IQ") + 
  ylab("UFO Observations (Per 100K Population)") + 
  theme_bw()
```

## Introduction Pieces (Data at a glance)

### Plotting the map of cities 

Use the first occurance of each city's coordinates, as they will be close enough anyway. 


```{r}
df 
```


```{r}
distinct_cities <- df %>% 
  filter(state %in% state.abb) %>% 
  distinct(city, state, .keep_all=TRUE)

distinct_cities <- distinct_cities %>% select("city", "state", "city_latitude", "city_longitude")
distinct_cities
```
```{r}
df
```


```{r}
states <- map_data("state")
```

```{r}
city_counts <- df %>%
  filter(state %in% state.abb) %>%
  group_by(state,city) %>% 
  summarize(count = n()) %>% 
  left_join(distinct_cities, by=c("state", "city"))
```
```{r}
city_counts_filtered <- city_counts %>% filter(count > 10)
```

```{r}
LA <- map_data("state", region="louisiana")
```

```{r}
MainStates <- map_data("state")

ggplot() + 
  geom_polygon(data=MainStates, aes(x=long, y=lat, group=group),
              color="black", fill="#D8DADC" ) + 
              geom_polygon() +
              coord_map() +
              geom_point(data=city_counts_filtered, aes(x=city_longitude, y=city_latitude, size=count), shape=21, colour="darkslateblue", fill="darkslateblue", alpha=0.8) + 
              scale_y_continuous(limits = c(25, 50)) + 
              scale_x_continuous(limits = c(-125, -60)) +
  labs(title="Occurances of UFO Sightings across U.S Cities 2015-2019",
       subtitle="represented for cities with at least 10 sightings. Size of points corresponds to sighting count") + 
  
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), 
        plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        )

ggsave("sighting_map.png")
```


### Basic Time Series


```{r}
trends %>% ggplot(aes(x=month, y=count, group=1)) + geom_line() +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  labs(title="UFO Sightings by Month (2015-2019)") + 
  xlab("Year-Month") + 
  ylab("Number of Sightings") + 
  scale_x_discrete(breaks=skip_four, guide = guide_axis(angle = 45)) +
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold",hjust = 0.5),
        text = element_text(size = 10),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 8))  
  

ggsave("month_sightings.png")
```


