---
title: "MADAR Exploration"
author: "Rio"
---
```{r}
library(tidyverse)

library(plotly)
```

```{r}
ufo_df <- read_csv("UFO_and_Weather.csv")
```
Remove non-US observations

```{r}
ufo_df<- subset(ufo_df, !(state %in% c("BC", "QC", "ON", "NS", "AB", "MB", "SK", "YT", "NL")))
```

Creates new variable 'contains_madar' that is a boolean (True/False) of whether the 'text' variable contains "madar" 
If it does, then this observation is an anomaly detection from a MADAR node, not a human UFO sighting 

```{r}
ufo_df_madar <- ufo_df
contains_madar <- grepl('MADAR', ufo_df_madar$text)

ufo_df_madar <- cbind(ufo_df_madar, contains_madar)
```

630 observations were from MADAR nodes while 21,245 observations were from human reports.
2.97% of all observations were from MADAR nodes.

```{r}
ufo_df_madar %>% 
  group_by(contains_madar) %>% 
  summarise(count=n())
```

the number of unique MADAR nodes (some nodes report multiple times)

```{r}
ufo_df_madar %>% 
  filter(contains_madar == TRUE) %>% 
  summarise(num_distinct_nodes = n_distinct(text))
```


Finds the number of MADAR node reports in each state.
Indiana leads with 99 reports followed by CT, Fl, VA, NH with ~ half of that 

```{r}
ufo_df_madar_by_state <- ufo_df_madar %>% 
  filter(contains_madar == TRUE) %>% 
  group_by(state) %>%
  summarise(node_report_count = n()) %>% 
  arrange(desc(node_report_count))

ufo_df_madar_by_state
```
Bar graph visualization of the number of reports from MADAR Nodes for each state

```{r}
ggplot(data = ufo_df_madar_by_state) +
  geom_bar(mapping = aes(x = reorder(state, -node_report_count), y = node_report_count), stat = "identity", fill = "#76A422") +
  labs(title="# of UFO recordings by MADAR Nodes by State", 
            y="Number of Reports from MADAR Nodes", 
            x="State") +
  theme_minimal()
```

Visualization on the US Map of the number of reports from MADAR Nodes 

```{r}
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
fig <- plot_geo(ufo_df_madar_by_state, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~node_report_count, text = ~node_report_count, locations = ~state,
    color = ~node_report_count, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "# of Reports from MADAR nodes")
fig <- fig %>% layout(
    title = 'UFO recordings by MADAR Nodes in the US from 2015-2019',
    geo = g
  )
fig
```
total number of UFO sightings (includes both human reports and MADAR node reports) for each state 

```{r}
ufo_sightings_by_state <- ufo_df_madar %>% 
  group_by(state) %>% 
  summarise(sightings = n()) %>% 
  arrange(desc(sightings))

ufo_sightings_by_state 
```
Whereas above we found the raw count of node reports by state, here we find the percentage of all reports for a state that the node reports constitute (so # node reports / # of all reports for each state)

We see that Indiana still leads at 22% followed by NH and CT both above 10% 

```{r}
ufo_madar_and_sightings <- merge(x=ufo_df_madar_by_state, y=ufo_sightings_by_state , by="state")
ufo_madar_and_sightings <- ufo_madar_and_sightings %>% 
  mutate(percent_node = node_report_count / sightings * 100) %>% 
  arrange(desc(percent_node))

ufo_madar_and_sightings
```
visualizes the above data of percent node reports by state 

```{r}
ggplot(data = ufo_madar_and_sightings) +
  geom_bar(mapping = aes(x = reorder(state, -percent_node), y = percent_node), stat = "identity", fill = "#76A422") +
  labs(title="% of Sightings Reported by MARDAR Nodes by State", 
            y="Percent of Sightings Reported by MARDAR Nodes", 
            x="State") +
  theme_minimal()
```

```{r}
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
fig <- plot_geo(ufo_madar_and_sightings, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~percent_node, text = ~percent_node, locations = ~state,
    color = ~percent_node, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "% of MADAR node Recordings")
fig <- fig %>% layout(
    title = '% of Recordings by MADAR Nodes in the US from 2015-2019',
    geo = g
  )
fig
```
Indiana has the most nodes located in the state. 

```{r}
ufo_madar_distinct_nodes_by_state <- ufo_df_madar %>% 
  filter(contains_madar == TRUE) %>% 
  group_by(state) %>% 
  summarise(num_distinct_nodes = n_distinct(text)) %>% 
  arrange(desc(num_distinct_nodes))

ufo_madar_distinct_nodes_by_state
```

Display both the number of distinct nodes and the number of node reports and percentage of reports from nodes by state

```{r}
ufo_madar_merge = merge(x = ufo_madar_distinct_nodes_by_state, y = ufo_madar_and_sightings, by = "state")
ufo_madar_merge %>% 
  arrange(desc(num_distinct_nodes))
```
Correlation of the number of distinct nodes to the number of reports from nodes 

```{r}
cor(ufo_madar_merge$num_distinct_nodes, ufo_madar_merge$node_report_count, use="complete.obs")
```

Correlation of the number of distinct nodes to the percentage of reports from nodes 

```{r}
cor(ufo_madar_merge$num_distinct_nodes, ufo_madar_merge$percent_node, use="complete.obs")
```
Plot of the correlation between number of distinct nodes and number of node reports

```{r}
ggplot(ufo_madar_merge, aes(x = num_distinct_nodes, y = node_report_count)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x) + 
  labs(title="Number of Node Reports vs Number of Distinct Nodes \n(Correlation = .86)", y="# Node Reports", x="# Distinct Nodes") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust=0.5)) +geom_text(x=750, y=350, label="Correlation=.68", size = 4)
ggsave("Number_Node_Reports_VS_Num_Distinct_Nodes.jpg", bg="white")
```
