---
title: "Weather Exploration"
author: "Rio"
---
```{r}
library(tidyverse)
library(plotly)
library(lubridate)
library(scales) 
```

```{r}
ufo_weather <- read_csv("UFO_and_Weather.csv")
```
Remove non-US observations

```{r}
ufo_weather <- subset(ufo_weather, !(state %in% c("BC", "QC", "ON", "NS", "AB", "MB", "SK", "YT", "NL")))
```

creates new variable 'weather_na_count' that counts the number of NAs from the weather variables 

```{r}
ufo_count_nas_weather <- ufo_weather %>%  
  mutate(weather_na_count = rowSums(is.na(select(., c(temperature, relative_humidity, precipitation, snow, wind_direction, wind_speed)))))

ufo_count_nas_weather
```
remove the observations that are completely missing all weather data (all 6 weather variables)
5,035 observations removed

```{r}
ufo_weather <- ufo_count_nas_weather %>% 
  filter(weather_na_count < 6)

ufo_weather
```
Notice that all observations contain NA for 'snow' 
Temperature appears to be the most complete variable with only 65 NAs compared to way more NAs for the other weather variables 

```{r}
summary(ufo_weather) 
```
Drop NAs for temperature to further analyze temperature 

```{r}
ufo_weather_dropna_tmp <- ufo_weather %>% 
  drop_na(temperature)
```


Temperature distribution from UFO sightings seems to be slightly negatively skewed

```{r}
ggplot(ufo_weather_dropna_tmp , aes(x = temperature)) +
      geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
      geom_density(alpha = .2, fill="#FF6655") 
```
The temperature during UFO sightings seems to align with seasonal patterns across the years. 

```{r}
ggplot(data = ufo_weather_dropna_tmp, aes(x=date_time, y=temperature)) +
  geom_point(na.rm=TRUE, color="darkolivegreen4", size=1, pch=20) +
  scale_y_continuous(breaks = seq(-50, 60, 10), limits=c(-30, 50)) +
  scale_x_datetime(breaks=date_breaks("1 year"), labels = date_format("%b %Y")) + 
  ggtitle("Air Temperature at Reported UFO Sighting Sites 2015-2019") +
  xlab("Date") + 
  ylab("Temperature (Celsius)") +
  theme_minimal()

ggsave("Air_Tmp_at_reported_UFO_Sighting_sites_2015_2019.jpg", bg="white")
```

find the average temperature during ufo sightings for each state 

```{r}
ufo_temp_by_state <- ufo_weather_dropna_tmp %>% 
  group_by(state) %>% 
  summarise(avg_tmp = mean(temperature))

ufo_temp_by_state
```
The average temperature during UFO sightings for each state plotted on a US map.

The average temperature during UFO sightings seems to align with state temperature trends where southern US is generally warmer than northern US. 
```{r}
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
fig <- plot_geo(ufo_temp_by_state, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~avg_tmp, text = ~avg_tmp, locations = ~state,
    color = ~avg_tmp, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Temperature")
fig <- fig %>% layout(
    title = 'Average Temperature of UFO Sightings in the US from 2015-2019',
    geo = g
  )
fig
```

A fairly uniform distribution of the wind direction during UFO sightings. 

```{r}
ufo_weather_dropna_wd <- ufo_weather %>% 
  drop_na(wind_direction)

ggplot(ufo_weather_dropna_wd , aes(x = wind_direction)) +
      geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
      geom_density(alpha = .2, fill="#FF6655") 
```
A positively skewed wind speed distribution for UFO sightings. The vast majority of the time, the wind speed is low. 

```{r}
ufo_weather_dropna_ws <- ufo_weather %>% 
  drop_na(wind_speed)

ggplot(ufo_weather_dropna_ws , aes(x = wind_speed)) +
      geom_histogram(bins = 30, aes(y = ..density..), colour = "black", fill = "white") 
```
There is almost never any precipitation during UFO sightings. 

```{r}
ufo_weather_dropna_precip <- ufo_weather %>% 
  drop_na(precipitation)

ggplot(ufo_weather_dropna_precip , aes(x = precipitation)) +
      geom_histogram(bins = 30, aes(y = ..density..), colour = "black", fill = "white")
```
Almost a normal distribution of relative humidity during UFO sightings. 

```{r}
ufo_weather_dropna_rh <- ufo_weather %>% 
  drop_na(relative_humidity)

ggplot(ufo_weather_dropna_rh , aes(x = relative_humidity)) +
      geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
      geom_density(alpha = .2, fill="#FF6655") 
```

